Ornaments as indicators of Social change before and after European contact at Kiwulan, Northeastern Taiwan

### 2019-05-25 analysis/paper/paper.Rmd
updates

1. Overall summary of  Raw material, shape/type and counts of ornaments
There are `r sample_size` ornaments excavated from the 40 sampling squares, which accounts for `r sample_pro`% of the total number of ornaments found in the excavation section where they were from. Figure \@ref(fig:plot-all-ornaments-count) shows that the most common ornament is golden bead with the number of 143, followed by 121 glass beads, 46 agate beads, 35 metal rings, and 22 metal bells. Ornaments in some categories are divided into sub-types due to the variations of shape, such as agate bead, metal ring, and metal bell. Table \@ref(tab:tab-all) shows the subtypes for the major ornaments.

To investigate whether the types of ornaments changes resulted from culture contact and how they changed, we compare ornaments from pre-European, European contact period, and Chinese contact period. The results a general pattern that most ornaments appeared in the pre-European contact period, reached a peak in quantity during the European contact and dropped down during the Chinese contact. This trend can be seen on the ornaments including golden bead, agate bead, metal ring, and bell. However, glass bead shows different pattern that higher frequency in the pre-European contact, and then the quantity goes down over time in the European and Chinese contact periods.

@@

```{r taball}
# clean some 
ornaments_period_clean <- 
ornaments_period %>% 
  mutate(Type = str_remove(Type, "prism/"),
         Type = ifelse(Type == "glass bead", "medium bead", Type))

ornaments_shape <-
  ornaments_period_clean %>% 
  group_by(period, Categories, Type ) %>% 
  tally() %>% 
  spread(period, n, fill = 0) %>% 
  filter(!Categories %in% c("Unknown Metal",

                            "Perforated plate", 
                            "Knited item",
                            "Hair pin",
                            "Bone Ornament",
                            "Metal Piece",
                            "Glass ring"))

@@

Table \@ref(tab:taball) shows the subtypes for the major ornaments.

%>% 
  spread(Categories, value)

 filter(Categories != "Unknown Metal") %>% 
%>% 
  mutate(Categories = factor(Categories,
                         level = c("Agate bead", 
                                 "European Contact", 
                                 "Chinese Contact"))) %>%

>>>>

1. Overall summary of  Raw material, shape/type and counts of ornaments
There are `r sample_size` ornaments excavated from the 40 sampling squares, which accounts for `r sample_pro`% of the total number of ornaments found in the excavation section where they were from. Figure \@ref(fig:plot-all-ornaments-count) shows that the most common ornament is golden bead with the number of 143, followed by 121 glass beads, 46 agate beads, 35 metal rings, and 22 metal bells. Ornaments in some categories are divided into sub-types due to the variations of shape, such as agate bead, metal ring, and metal bell. Table \@ref(tab:tab-all) shows the subtypes for the major ornaments. 
## Comparisons between time periods

To investigate whether the types of ornaments changes resulted from culture contact and how they changed, we compared ornaments from pre-European, European contact period, and Chinese contact period. The results show a general pattern that most ornaments appeared in the pre-European contact period, their frequencies reach a peak during the European contact and dropped significantly in the Chinese contact period, especially for golden bead. This trend can be also seen on ornaments including agate bead, metal ring, and bell. However, glass bead shows a different pattern that indicates a higher frequency in the pre-European contact, and then the frequency decreased in the European contact period and even less in the Chinese contact period.

To further understand whether there are some changes in the frequency of subtypes mainly defined by the shape, we compared the distribution of frequency for each major categories. For the

For golden bead, we focus on their length that

@@

```{r taball}
# clean some 
ornaments_period_clean <- 
ornaments_period %>% 
  mutate(Type = str_remove(Type, "prism/"),
         Type = ifelse(Type == "glass bead", "medium bead", Type),
         Type = ifelse(Type == "small I", "plain small", Type),
         Type = ifelse(Type == "small II", "thin small", Type))

ornaments_shape <-
  ornaments_period_clean %>% 
  group_by(period, Categories, Type ) %>% 
  tally() %>% 
  spread(period, n, fill = 0) %>% 
  filter(!Categories %in% c("Unknown Metal", 
                            "Perforated plate", 
                            "Knited item",
                            "Hair pin",
                            "Bone Ornament",
                            "Metal Piece",
                            "Glass ring"))

@@
